Iroko: Predictive Datacenter Congestion Control

### 2017-12-10 Documents/Report/Iroko_Report.tex
More graphs
report updated
more changes to intro

\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}

@@

\begin{abstract}
Datacenter networks have become a hotbed for research in the recent past. Many works have been published on improving bisection bandwidth, cost, and latency. As a core contribution of this research, centralized scheduling has entered the stage as a dominant strategy to manage flows. Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization. However, modern congestion control algorithms and schedulers are still fundamentally reactive. Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler. Using tokens as a global tool to enforce traffic limits and admission control, the scheduler is able to proactively steer the flow of traffic. The notion of "knowledge" and predictive analysis in networks is a growing trend in research, which we intend to leverage in this system.

We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions. Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks. We will compare our "Iroko" system against established TCP-congestion systems such as Hedera and DCTCP. Based on the findings and measurement results, we will reassess the feasibility and prospects of a predictive congestion control algorithm.
\end{abstract}

@@

\input{intro}
%\input{relatedwork}
\input{design}
% %\input{motivation}
\input{implementation}
\input{eval}

\bibliographystyle{ACM-Reference-Format}
\bibliography{Iroko_Proposal}

>>>>

\title{Iroko: Predictive Datacenter Congestion Control}

@@

\begin{abstract}
Datacenter networks have become a hotbed of network research in the past 
decade. Many works have been published on improving bisection bandwidth, cost, 
and latency. As a core contribution of this research, centralized scheduling 
has entered the stage as a dominant strategy to manage flows. Benefitting from 
a global view and full control over the network, centralized schedulers are 
able to enforce fine-grained traffic control, frequently achieving near-optimal 
bandwidth optimization. However, modern congestion control algorithms and 
schedulers are still fundamentally reactive. Many algorithms are designed to 
only respond to packet loss or significant increase in latency, not to actively 
prevent it.
In this project, we plan to explore the opportunities of employing a preemptive 
and tightly controlling central network scheduler. By enforcing traffic limits 
and admission control, the scheduler is able to proactively steer the flow of 
traffic. The notion of "knowledge" and predictive analysis in networks is a 
growing trend in research, which we intend to leverage in this system.

We investigate the potential in such a new form of interactive congestion 
control and analyze it against state-of-the-art solutions. Our tool of choice 
to model our system is Mininet, a rapid prototyping emulator for datacenter 
networks. We will compare our "Iroko" system against established TCP-congestion 
systems such as Hedera and DCTCP. Based on the findings, measurement results, 
and effectiveness of our reinforcement learning technique, we will reassess the 
feasibility and prospects of a predictive congestion control algorithm.
\end{abstract}

@@

\input{intro}
%\input{motivation}
\input{design}

\input{implementation}
\input{eval}
\input{discussion}
\input{future_work}
\input{conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{Iroko_Report}

### 2017-12-10 Documents/Report/intro.tex
more changes to intro

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, data centers must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.
Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
is a perennial challenge for any network operator.
Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the 
eventual collapse of a high-load network, requiring sophisticated 
approaches to solve congestion control.
Despite the innovation potential of data centers, TCP has dominated as 
the congestion control protocol of choice.
TCP is a \textit{reactive protocol}, responding to indicators of congestion and 
latency in the network. However, the fact that packet loss and latency 
surges occur in the network, already portends a problem. Packet loss is 
incurred by overflowing queues in forwarding elements or mismatched 
hardware capabilities, implying that traffic has not been optimally 
distributed. Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

@@

Moving away from the principle of distributed communication and routing, SDN 
introduces the notion of "centralized management". A single controller with 
global knowledge is able to automatically modify and adapt the forwarding 
tables of all switches in the network, while notifying end hosts of changes in 
the network.
These two new trends in system design; full architectural control and 
centralized management, facilitated new opportunities in the space of TCP 
congestion research. Traffic can now be managed in a  \textit{centralized} 
fashion based on \textit{global knowledge} of the entire topology and traffic 
patterns.

A new line of centralized schedulers has emerged that can achieve close to 
optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
However, these schedulers are still  \textit{reactive}  in nature. The central 
controller responds to changes in the network or requests by applications, 
which may cost valuable round-trip latency. Often, short-term flows or bursts 
are unaccounted for, which causes undesirable packet loss and back-propagating 
congestion.

The idea of admission control and service guarantees in networks is not 
new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a datacenter, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behavior, which allows for great simplification of enforcement 
and prioritization policies.

A desirable solution is a global, centralized arbiter which is able to predict 
and fairly distribute flows in the network before bursts or congestion even 
occur. By treating the network's compute and forwarding power as a single 
finite resource, a controller acts akin to a OS scheduler distributing CPU time 
slices to processes. This design approach follows SDN's aspiration of 
introducing operating systems abstractions to the networking domain space.

In this project, we plan to explore the possibilities of a centralized, 
proactive flow scheduler. We ask ourselves the following research questions:
\begin{enumerate}
    \item What are the requirements for such a centralized, predictive 
    scheduler to succeed?
    \item Is it possible to preemptively regulate a network by analyzing global 
    traffic patterns?
    \item What latency, packet loss and utilization are we able to achieve?
\end{enumerate}

In the scope of this course, we attempt to answer these questions and design a 
simple predictive scheduler in Mininet. If successful, we will benchmark our 
results and evaluate the level of utilization compared to contemporary 
scheduling systems.

>>>>

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, data centers must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.
Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
is a perennial challenge for any network operator.

Inefficient routing and congestion control can quickly lead to 
bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load 
network, requiring more sophisticated solutions. In fact, recent insights into 
datacenter network traffic have shown that buffer overruns and small 
short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc}
Despite the innovation potential of data centers, TCP has dominated as 
the congestion control algorithm of choice.
TCP is a \textit{reactive protocol}, responding to indicators of packet drops
latency. However, this surge of loss and latency in the network already 
portends a problem. Overflowing queues in forwarding elements or mismatched 
hardware capabilities imply that traffic has not been optimally managed.
Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

While TCP has not fundamentally changed in

@@

Moving away from the principle of distributed communication and routing, SDN 
introduces the notion of "centralized management". A single controller with 
global knowledge is able to automatically modify and adapt the forwarding 
tables of all switches in the network, while notifying end hosts of changes in 
the network.
This full architectural control facilitated new opportunities in the space of 
TCP congestion research. Traffic can now be managed in a  \textit{centralized} 
fashion based on \textit{global knowledge} of the entire topology and traffic 
patterns.

Following SDN's rise in popularity, a new line of centralized schedulers that 
can achieve close to optimal bandwidth utilization has emerged.~\cite{hedera, 
fastpass, microte, b4, dionysus}
However, these schedulers are still \textit{reactive} in nature. The central 
controller responds to changes in the network or requests by applications, 
which may cost valuable round-trip latency. Often, short-term flows or bursts 
are unaccounted for, which causes undesirable packet loss and back-propagating 
congestion.~\cite{perc}

In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
While idea of admission control and service guarantees in networks is not 
new,~\cite{access_limit, access_limit2} these designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a data center, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies. Admission control systems can give administrators 
the benefit of tight control over the data center, mitigating the effects of 
queue-buildup and bursts.

Although it is unclear how predictable data center traffic truly is, various 
works have shown that repeated patterns may 
exist.\cite{msr_dc,fb_dc,traffic,microte}
If data center traffic exhibits predictable patterns over long time, it may be 
possible to devise a proactive mechanism which predicts the traffic matrix of 
the next nth iterations, and adjusts traffic accordingly.

%Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is possible to 
%learn optimal TCP congestion control parameters for various network 
%topologies. 
%Benson et al.,~2014\cite{microte} analyzed data center traffic and inferred 
%that around 30\%

An ideal culmination of these insights may be a global, 
centralized arbiter, able to predict and fairly distribute flows in the 
network. By only allowing a safe sending rate for each participant it minimizes 
the amount of bursts and loss possible.
Treating the network's compute and forwarding power as a single finite 
resource, the controller acts akin to a OS scheduler distributing CPU time 
slices to processes.

This project report details our exploration of such a centralized, 
proactive flow scheduler. We asked ourselves the following research questions:
\begin{enumerate}
    \item What are the requirements for such a centralized, predictive 
    scheduler to succeed?
    \item Is it possible to preemptively regulate a network by analyzing global 
    traffic patterns?
    \item What latency, packet loss and utilization are we able to achieve?
\end{enumerate}
In the scope of this report, we outline the design and implementation of our 
flow scheduler and learning approach. We perform an analysis of its 
effectiveness compared to DCTCP and Hedera and conclude with a discussion of 
the benefits and drawbacks of centralized learning and admission control in the 
data center context.
